Constant-round Blind Classical Verification of Quantum Sampling

### 2021-10-22 introduction.tex
Remove comments from intro

Recently, the problem is resolved by a breakthrough result of Mahadev~\cite{FOCS:Mahadev18a}, who constructed the first Classical Verification of Quantum Computation (CVQC) protocol for $\BQP$, where an efficient classical ($\BPP$) verifier can interact with an efficient quantum ($\BQP$) prover to verify any $\BQP$ language. Soundness of Mahadev's protocol  is based on a widely recognized computational assumption that the learning with errors (LWE) problem~\cite{JACM:Regev09} is hard for $\BQP$ machines.
%In a recent breakthrough, Mahadev~\cite{FOCS:Mahadev18a} proposed the first protocol of Classical Verification of Quantum Computation (CVQC) whose soundness is based on a widely recognized computational assumption that the learning with error (LWE) problem~\cite{JACM:Regev09} is hard for $\BQP$ machines.
The technique invented therein has inspired many subsequent developments of CVQC protocols with improved parameters and functionality. For example, Mahadev's protocol has a large constant soundness error. The works of ~\cite{arXiv:AlaChiHun19,arXiv:ChiaChungYam19} use parallel repetition to achieve a negligible soundness error. As another example, the work of ~\cite{FOCS:GheVid19} extends Mahadev's techniques in an involved way to obtain a CVQC protocol with an additional blindness property.

%For example,~\cite{arXiv:AlaChiHun19,arXiv:ChiaChungYam19}  
%
%(e.g.,~\cite{FOCS:GheVid19,arXiv:AlaChiHun19,arXiv:ChiaChungYam19}).

@@

% On the other side, there are known complexity-theoretic limitations on the feasibility of blind CVQC in the information-theoretical setting (e.g.~\cite{aaronson_et_al:LIPIcs:2019:10582}).

% With the newly developed techniques, we revisit the classical verification of quantum computation problems from both a philosophical and a practical point of view.

@@

We initiate the study of CVQC for quantum sampling problem, which we believe is highly desirable and natural for delegation of quantum computation. Due to the intrinsic randomness of quantum mechanics, the output from a quantum computation is randomized and described by a distribution. 
% In general, a round of quantum computation can be described as applying a quantum circuit and measuring the output qubits in computational basis.
%To verifiably delegate general quantum computation, it is natural to require that the output is close to the correct distribution when the verifier accepts. 
Thus, if a classical verifier want to utilize the full power of a quantum machine, the ability to get a verifiable sample from the quantum circuit's output distribution is desirable.  On a more concrete level, quantum algorithms like Shor's algorithm~\cite{Shor} has a significant quantum sampling component, and the recent quantum supremacy proposals (e.g.,~\cite{Boson, IQP, nature-google}) are built around sampling tasks, suggesting the importance of sampling in quantum computation.

%The previous works following Madadev's construction has been focusing on decision problems. In this work, we proposed and constructed a extension of Mahadev's protocol to \emph{sampling} problems, where the classical verifier get a sample from a desired distribution from the quantum prover. More formally, we seek a classical verifiable delegation of the class $\SampBQP$ proposed by Aaronson~\cite{aaronson_2013}.  Such an extension to sampling problems is desirable on a fundamental level; since we can always measure all qubits at the end of a quantum circuit, every quantum circuit naturally gives some distribution to sample from. Thus, if a classical verifier want to utilize the full power of a quantum machine, the ability to get a verifiable sample from the quantum circuit's output distribution is desired.  On a more concrete level, quantum algorithms like Shor's algorithm~\cite{Shor} has a significant quantum sampling component, and the recent quantum supremacy proposals (e.g.,~\cite{Boson, IQP, nature-google}) are built around sampling tasks, suggesting the importance of sampling in quantum computation.

% We first observe that the \emph{sampling} version of $\BQP$ (e.g., the class $\SampBQP$ formulated by Aaronson~\cite{aaronson_2013}) might be a more appropriate notion to serve the purpose of the original problem.

% Moreover, a lot of quantum algorithms from Shor's~\cite{Shor} and Grover's~\cite{Grover} algorithms to some recent developments in machine learning and optimization (e.g.~\cite{brando_et_al:LIPIcs:2019:10603, AGGW17,pmlr-v97-li19b}) contain a significant quantum sampling component.
% The fact that almost all quantum supremacy tasks (e.g.,~\cite{Boson, IQP, nature-google}) are sampling ones also suggests the relevance of delegation for quantum sampling problems.

It is worth noting that the difficulty of extending the delegation of decision problem to the delegation of sampling problems is quantum-specific. This is because 
there is a simple reduction from the delegation of \emph{classical} sampling problems to decision ones:  the verifier can sample and fix the random seed of the computation, 
%\footnote{One can use pseudorandomness, e.g., generated by a PRG seed, if verifier's efficiency is required, as typically in the classical delegation of computation literature.}, 
which makes the computation deterministic. Then, the verifier can delegate the output of the computation bit-by-bit as decision problems. However, this derandomization trick does not work in the quantum setting due to its intrinsic randomness.

%fix a seed for pseudo-randomness, send it to the prover, and then the sampling outcome can be computed  bit-by-bit under this pseudo-random seed in a deterministic way. Therefore, classical literature of delegation of computation primarily focuses on the delegation of decision problems. 
%However, it is unclear how to make this de-randomization trick work for the delegation of $\SampBQP$,  as randomness in quantum computation comes from inherent nature of quantum mechanics rather than some random seeds. 
%Thus, it seems that the delegation of $\SampBQP$ needs to take a different technical route.

% The first and important step in the case of $\SampBQP$ is to identify an appropriate definition of corresponding CVQC protocols.

@@

\paragraph{Our Contribution.}
As the first step to formalize CVQC for quantum sampling problems, we consider the complexity class $\SampBQP$ introduced by Aaronson~\cite{aaronson_2013} as a natural class to capture efficiently computable quantum sampling problems. 
%model
%formalize efficiently computable quantum sampling problems with the complexity class $\SampBQP$ introduced by Aaronson~\cite{aaronson_2013}.
%
%Formally, 
$\SampBQP$ consists of sampling problems $(D_x)_{x\in\zo^*}$ that can be approximately sampled by a $\BQP$ machine with a desired inverse polynomial error (See Section~\ref{sec:samp_definition} for the formal definition). We consider CVQC for a $\SampBQP$ problem $(D_x)_{x\in\zo^*}$ where a classical $\BPP$ verifier delegates the computation of a sample $z\leftarrow D_x$ for some input $x$ to a quantum $\BQP$ prover. Completeness requires that when the prover is honest, the verifier should accept with high probability and learn a correct sample $z\leftarrow D_x$. For soundness, intuitively, the verifier should not accept and output a sample with incorrect distribution  when interacting with a malicious prover. We formalize the soundness by a strong \emph{simulation-based} definition, (\Cref{dfn:stats-secure-proto-sampbqp}),
where we require that the joint distribution $(d,z)$ of the decision bit $d \in \set{\Acc, \Rej}$ and the output $z$ (which is $\bot$ when $d = \Rej$) is $\eps$-close (in  either statistical or computational sense) to an ``ideal distribution'' $(d,z_{ideal})$, where $z_{ideal}$ is sampled from the desired distribution $D_x$ when $d = \Acc$ and set to $\bot$ when $d = \Rej$.\footnote{This simulation-based formulation is analogous to the standard composable security definition for QKD.} \hannote{justify?}
%
%To the best of our knowledge, we are the first to formally define delegation of quantum sampling problems.

%and almost never accept and output from an incorrect distribution when interacting with a malicious prover. We formalize this by a strong \emph{simulation-based} definition (Section~\ref{sec:samp_definition}), where we require that the joint distribution of the decision bit $d \in \set{\Acc, \Rej}$ and the output $z$ (which is $\bot$ when $d = \Rej$) is $\eps$-close (in  either statistical or computational sense) to an ``ideal distribution'' $(d,z_{ideal})$, where $z_{ideal}$ is sampled from the desired distribution $D_x$ when $d = \Acc$ and set to $\bot$ when $d = \Rej$.

@@

%, which we discuss  further in Section~\ref{subsec:discussion}.

%\KM{mention the inverse poly error and add forward pointer for discussion}

% To the best of our knowledge, we are the first to formally define the delegation of quantum sampling problems.

% We note that while several constructions in the relaxed models (e.g., verifiable blind computation~\cite{FK17}) can be naturally generalized to delegate quantum sampling problems and allow the verifier to learn multi-bit outputs, it seems non-trivial to show that these constructions achieve the simulation-based soundness property we defined. Proving the soundness of these constructions for delegating $\SampBQP$ is an interesting open question.

% The requirement that the verifier needs to output a sample when accepts makes a direct application of Mahadev's protocol infeasible for $\SampBQP$. As a result, in order to construct a $\QPIP_0$ protocol for $\SampBQP$,  one needs to (1) construct a $\QPIP_1$ protocol for $\SampBQP$ with very small completeness and soundness errors, and (2) amend Mahadev's construction so that the protocol will generate the desired sample whenever it accepts. 
% Our contribution is a solution to address the above two technical challenges.

% In contrast to the single-bit $\mathrm{Accept/Reject}$ information for the decision $\BQP$, any $\SampBQP$ problem needs to always output a sample of the desired distribution.
% In the context of CVQC protocols, it means that the protocol should allow the verifier to generate the desired output sample whenever the protocol accepts. Of course, when the prover cheats, the verifier will reject and no further output is required in that case.  
% Intuitively, the soundness should require that the verifier $V$ should never be ``cheated'' to accept and output an incorrect sample even when interacting with a malicious prover.

% 
% Indeed, some nice features of the Hadamard round,  e.g., allowing X-Z measurements on any qubit,  suggest the feasibility of generating  multi-bit measurement outcomes with the protocol.

%\paragraph{Techniques.}

@@

%Our protocol is constructed by generalizing Mahadev's protocol to ha

\paragraph{Overview of Mahadev's Protocol.}
%
Following~\cite{FOCS:Mahadev18a}, we define $\QPIP_{\tau}$ as classes of interactive proof systems between an (almost) classical verifier and a quantum prover, where the classical verifier has limited quantum computational capability, formalized as possessing $\tau$-qubit quantum memory. %\footnote{Intuitively, the size of local quantum memory limits the size of qubits that a general (entangled) quantum operation can be operated on at the same time. However, the verifier is still able to operate on all the qubits in a streaming fashion: i.e., receive the qubits sent from the prover and discard some local qubits in the case of overflow.}. 
A formal definition is given in Appendix~\ref{sec:qpip_def}.

@@

\paragraph{Overview of Our Construction.}
%As mentioned, our construction of CVQC protocol for $\SampBQP$ 
Following the blueprint of Mahadev's construction, our construction proceeds in the following two steps: 1. construct a $\QPIP_1$ protocol for $\SampBQP$ with required special property, and 2. compile the $\QPIP_1$ protocol using $\PiMeasure$ to get the desired $\QPIP_0$ protocol. The first step can be done by combining existing techniques from different contexts, whereas the second step is the main technical challenge.
%Both steps requires non-trivial works, but it turns out that the second step is the main technical challenge. 
At a high-level, the reason is the above-mentioned issue that the verifier does not learn the measurement outcome in the testing round. While this is not a problem for decision problems, %(where the verifier only needs to decide to accept or reject), 
for sampling problems, the verifier needs to produce an output sample when accepts, but there seems to be no way to produce the output for the verifier without learning the measurement outcome.   
%
We discuss both steps in turn as follows.

%As we shall see, both steps requires non-trivial works, and the second step is the main technical challenge.

@@

Clearly from the construction, the verifier now can output a sample when it decides to accept, and output a correct sample when interacting with an honest prover (completeness). The challenge is to show that $\PiSampZ$ is computationally sound. Since we are now in the computational setting, we cannot use the quantum de Finetti theorem as above which only holds in the information-theoretical setting\hannote{really? maybe emphasis on constructing adversary}. Furthermore, parallel repetition for computationally sound protocols are typically difficult to analyze, and known to not always work for protocols with four or more messages even in the classical setting~\cite{BIN97,PW12}.

% Fortunately, parallel repetition of Mahadev's protocol for $\BQP$ has been analyzed before in ~\cite{arXiv:ChiaChungYam19, arXiv:AlaChiHun19} with interestingly different proofs. Both works showed that $m$-fold parallel repetition reduces the soundness error to $2^{-m}$ for $\BQP$ by relying on special properties of Mahadev's protocol.

 Parallel repetition of Mahadev's protocol for $\BQP$ has been analyzed before in ~\cite{arXiv:ChiaChungYam19, arXiv:AlaChiHun19}. However, the situation here is different. 
 %the parallel repetitions of Mahadev's protocol for $\BQP$ and $\SampBQP$ is very different. 
 For $\BQP$, the verifier simply chooses to run the Hadamard and testing rounds independently for each repetition.
 In contrast, our $\PiSampZ$ runs the Hadamard round in one repetition and runs the testing rounds in the rest. The reason is that in $\SampBQP$, as well as generically in sampling problems, there is no known approach to combine multiple samples to generate one sample with reduced error, i.e., there is no generic error reduction method for the sampling problem. 
In contrast, the error reduction for decision problems can be done with the majority vote. 
As a result, while the soundness error decreases exponentially for $\BQP$, as we see below (and also in the above $\QPIP_1$ protocols), for $\SampBQP$, $m$-fold repetition only decreases the error to $\poly(1/m)$.

%Intuitively, Mahadev's protocol has the property that if a malicious prover $P^*$ knows the answer to pass the testing round, then it must be ``committed'' to a state $\rho$ in the sense that the verifier $V$ learns the $X/Z$ measurement outcome of $\rho$ in the Hadamard round (in a computational sense). Thus, the soundness of the underlying $\QPIP_1$ protocol ensures that $V$ learns a correct sample when it accepts. In short, when $P^*$ knows the testing round answer,  $P^*$ cannot ``cheat'' in the Hadamard round.

%\hannote{copy ch 5 here?}
To analyze the soundness of $\PiSampZ$, we use the \emph{partition lemma} developed in~\cite{arXiv:ChiaChungYam19} to analyze the prover's behavior while executing copies of $\PiMeasure$.\footnote{The analysis of \cite{arXiv:AlaChiHun19} is more tailored to the decision problems setting, and it is unclear how to extend it to sampling problems where there are multiple bits of output.} Intuitively, the partition lemma says that for any cheating prover and for each copy $i\in[m]$, there exist two efficient ``projectors" \footnote{Actually they are not projectors, but for the simplicity of this discussion let's assume they are.} $G_{0,i}$ and $G_{1,i}$ in the prover's internal space with $G_{0,i}+G_{1,i} \approx Id$. $G_{0,i}$ and $G_{1,i}$ splits up the prover's residual internal state after sending back his first message.
$G_{0,i}$ intuitively represents the subspace where the prover does not knows the answer to the testing round on the $i$-th copy, while $G_{1,i}$ represents the subspace where the prover does. Note that the prover is using a single internal space for all copies, and every $G_{0,i}$ and every $G_{1,i}$ is acting on this single internal space. 
By using this partition lemma iteratively, we can decompose the prover's internal state $\ket{\psi}$ into sum of subnormalized states.
First we apply it to the first copy, writing $\ket{\psi}=G_{0,1}\ket{\psi}+G_{1,1}\ket{\psi} \equiv \ket{\psi_0}+\ket{\psi_1}$.
The component $\ket{\psi_0}$ would then get rejected as long as the first copy is chosen as a testing round,
which occurs with pretty high probability.
More precisely, the output corresponding to $\ket{\psi_0}$ is $1/m$-close to the ideal distribution that just rejects all the time.
On the other hand, $\ket{\psi_1}$ is now binding on the first copy;
we now similarly apply the partition lemma of the second copy to $\ket{\psi_1}$.
We write $\ket{\psi_1}=G_{0,2}\ket{\psi_1}+G_{1,2}\ket{\psi_1}\equiv \ket{\psi_{10}}+\ket{\psi_{11}}$, and apply the same argument about $\ket{\psi_{10}}$ and $\ket{\psi_{11}}$.
We then continue to decompose $\ket{\psi_{11}}=\ket{\psi_{110}}+\ket{\psi_{111}}$ and so on, until we reach the last copy and obtain $\ket{\psi_{1^m}}$.
Intuitively, all the $\ket{\psi_{1\dots10}}$ terms will be rejected with high probability, while the $\ket{\psi_{1^m}}$ term represents the ``good" component where the prover knows the answer to every testing round and therefore has high accept probability. Therefore, $\ket{\psi_{1^m}}$ also satisfies some binding property,
so the verifier should obtain a measurement result of some state on the Hadamard round copy,
and the soundness of the $\QPIP_1$ protocol $\PiSamp$ follows.

@@

% ~\cite{arXiv:ChiaChungYam19}  introduces a \emph{partition lemma}, which roughly says that there is an efficient projection that partitions the prover's internal quantum space into a ``known-answer'' subspace $S_1$ where $P^*$ knows the answer to the testing round, and a ``not-known-answer'' subspace $S_0$ where $P^*$ does not know the answer to the testing round. For an (efficient) prover's states $\ket{\psi} \in S_1$, the verifier $V$ can produce a sound output in the Hadamard round. For (efficient) prover's states $\ket{\psi} \in S_0$, the verifier $V$ rejects in the testing round with high probability. The work of \cite{arXiv:ChiaChungYam19} used this partition lemma iteratively to each instance of the repetition to show that for a no instance $x$ for any (efficient) prover's state $\ket{\psi}$, the probability that $V$ accepts is at most roughly $2^{-m}$.

% We apply the partition lemma iteratively to the prover's state  similarly to~\cite{arXiv:ChiaChungYam19}, but with significant differences to show the soundness of $\PiSampZ$ for $\SampBQP$. In particular, extra care is required to reason about the computational indistinguishability of the verifier's output sample from the correct distribution $D_x$. To gain some intuition, let us apply the partition lemma to the prover's state $\ket{\psi}$ with respect to the first coordinate, which decompose $\ket{\psi} = \ket{\psi_0} + \ket{\psi_1}$ with $\ket{\psi_0} \in S_0$ and  $\ket{\psi_1} \in S_1$, respectively. Intuitively, the $\ket{\psi_0}$ component will be rejected whenever the first coordinate runs the testing round, but the verifier may be ``cheated'' to accepted incorrectly when the first coordinate runs the Hadamard round, which happens with probability $1/m$ and is where the error come from. For the $\ket{\psi_1}$ component, $V$ will produce a sound output if it runs the Hadamard round at the first coordinate. For the case that $V$ chooses to run the testing round at the first coordinate, we can analyze it by further applying the partition lemma to $\ket{\psi_1}$ with respect to the second coordinate, and repeat the analysis, and so on. 
% %
The intuitive analysis outlined above glosses over many technical details, and we substantiate this outline with full details in 
Section~\ref{sec:qpip0_all}.

%\hannote{put this somewhere else?}
%It is worthwhile mentioning that we came to notice some online discussion\footnote{\url{https://www.scottaaronson.com/blog/?p=3697}, e.g., comment \#25, \#26, \#42, \#48. } on the possibility of a CVQC protocol for $\SampBQP$ after we developed our own result. These comments suggested a possible reduction of $\SampBQP$ to the local Hamiltonian problem following a similar high-level idea as our solution, however, with no detail and a seemingly different technical route. More importantly, the issue that the verifier does not generate outputs in the testing round seems to be overlooked and not discussed there.

%A natural attempt to construct a CVQC protocol for $\SampBQP$ is to directly apply Mahadev's CVQC protocol for $\SampBQP$~\cite{FOCS:Mahadev18a}.  However, the requirement that the verifier needs to get a sample whenever accepting is not easily fulfilled by Mahadev's protocol. Before we explain this issue in details, let us go through a brief review of the construction and notations of Mahadev's protocol.
%\hannote{something more nature} \hannote{explain mahadev protocol} \hannote{explain how to direct apply} \hannote{qpip1}
% Let us revisit Mahadev's CVQC protocol~\cite{FOCS:Mahadev18a} first for some technical background. 
%Following~\cite{FOCS:Mahadev18a}, we formally define $\QPIP_{\tau}$

%Following~\cite{FOCS:Mahadev18a}, we define $\QPIP_{\tau}$ as classes of interactive proof systems between an (almost) classical verifier and a quantum prover, where the classical verifier has limited quantum computational capability, formalized as possessing $\tau$-qubit quantum memory\footnote{Intuitively, the size of local quantum memory limits the size of qubits that a general (entangled) quantum operation can be operated on at the same time. However, the verifier is still able to operate on all the qubits in a streaming fashion: i.e., receive the qubits sent from the prover and discard some local qubits in the case of overflow.}. 
%A formal definition is given in Section~\ref{sec:qpip_def}. 
%It is known that a solution to a $\BQP$-problem can be efficiently verified by a classical verifier who can perform a single qubit $X$ or $Z$ measurement~\cite{PhysRevA.93.022326, mf16}, by reducing any $\BQP$ problem to a local Hamiltonian problem where each term consists of  $X$ and $Z$ only. This leads to a $\QPIP_1$ protocol for $\BQP$.
%The main contribution of Mahadev~\cite{FOCS:Mahadev18a} is a compiler from a $\QPIP_1$ protocol with $X$ or $Z$ measurements to a $\QPIP_0$ protocol.
% ====
% In this section, we create a delegation protocol for $\SampBQP$ with fully classical clients by adapting the approach taken in \cite{FOCS:Mahadev18a}.  
%The center piece of Mahadev's compiler is the protocol $\PiMeasure$ (See \Cref{proto:urmila4}) that allows a $\BQP$ prover to ``commit a state" for a classical verifier to choose a $X$ or $Z$ measurement to measure on.
% Composing it with the $\QPIP_1$ protocol for $\BQP$ from \cite{mf16} results in a $\QPIP_0$ protocol for $\BQP$.
%$\PiMeasure$ is a 4-round protocol between a verifier  and a prover.
%The verifier start by secretly choosing a string $h$ specifying the $X$ or $Z$ measurements it wants to make on each qubit, and generates keys $pk, sk$ from $h$. The first round of communication is the verifier sending $pk$ to the prover. The prover ``commits" to a state $\rho$ of its choice using $pk$, and the second round of communication is the prover replying with its commitment $y$.
%The verifier must then choose between two options: a \emph{testing round} or a \emph{Hadamard round}.
%If the verifier chose the testing round, it can catch cheating provers at the end of the protocol, 
%and if the verifier chose the Hadamard round, it receives desired $XZ$-measurement outcomes at the end of the protocol.
%The third round of communication is the verifier sending his 1-bit choice of testing or Hadamard round to the prover. Finally, in the fourth round of the communication, the prover replies accord to the verifier's choice in the third round. After the last round of the communication, the verifier wrap up the protocol with some extra computation. If the testing round was chosen, the verifier checks the prover's fourh-round reply against the previous commitment, and rejects if he sees an inconsistency. If the Hadamard round was chosen, the verifier calculates $M_{XZ}(\rho, h)$, the $XZ$-measurement specified by $h$ on $\rho$, basing on the fourth-round reply. Using $\PiMeasure$ to delegate the $X$ or $Z$ measurement of the aforementioned $\QPIP_1$ protocol~\cite{mf16} results in a $\QPIP_0$ protocol for $\BQP$.

% In this work, we will compose $\PiMeasure$ with our $\QPIP_1$ protocol $\PiSamp$ (\Cref{ProtoQPIP1}) for $\SampBQP$ in order to obtain a $\QPIP_0$ protocol for $\SampBQP$.

% A direct composition of $\PiSamp$ and $\PiMeasure$, however, results in a protocol $\PiNaive$ (\Cref{proto:qpip0_naive}) which does not provide reasonable completeness or accuracy guarantees.
% This is due to $\PiMeasure$ itself having peculiar and weak guarantees:
% the client doesn't always obtain measurement outcomes even if the server were honest.

% When that happens under the $\BQP$ context, the verifier can simply accept the prover at the cost of some soundness error;
% under our $\SampBQP$ context, however, we must run many copies of $\PiNaive$ in parallel so the verifier can generate its outputs from some copy.
% We will spend the majority of this section analyzing the soundness of this parallel repetition.

% \hannote{4 round?} As we described in blah, Mahadev's compiler is a 4-round protocol. 
% However, the requirement that the verifier needs to get a sample whenever accepting exposes one drawback of Mahadev's protocol.

% Note that Mahadev's protocol consists of the so-called testing round, where the prover's behavior is tested, and the Hadamard round, where measurement outcomes are generated. 
% Mimicing Mahadev's construction of

%With Mahadev's compiler from $\QPIP_1$ to $\QPIP_0$ for $\BQP$ in mind, a natural approach to construct a $\QPIP_0$ protocol for $\SampBQP$ is the follow step approach: 1. construct a $\QPIP_1$ protocol for $\SampBQP$ with $X$ or $Z$ measurements. 2. compile the $\QPIP_1$ protocol with $\PiMeasure$ to get $\QPIP_0$ protocol. This indeed is the general outline we follow in this work. However, getting our result is not straightforward. While it is relatively straightforward to do the first step with techniques in the literature, a naive combination between the resulting $\QPIP_1$ protocol and $\PiMeasure$ does not give a $\QPIP_0$ protocols for $\SampBQP$ with reasonable \emph{completeness}. The reason is that while it is possible to generate desired output samples in the Hadamard round of $\PiMeasure$, 
%it is unclear how to generate such a sample when the protocol accepts in the testing round. 
%Note that this difficulty originates from the nature of sampling problems. In a decision problem, the verifier can blindly choose to output accept even if he did not get the measurement details, potential sacrificing some soundness. In comparison, in a sampling problem, the verifier always needs to output a good sample.

%\hannote{our solution}

%\hannote{something quick on qpip1, can be put together with available techs. scott blog }

%\hannote{japanese citations : \cite{takeuchi2018verification,hayashi2015verifiable}}

@@

%Mahadev's construction~\cite{mahadev_qfhe}.
%building on top of quantum fully homomorphic encyprtion (QFHE). 
%Precisely, we leverage QFHE (especially the one from~\cite{mahadev_qfhe}) to transform any CVQC protocol to \emph{a blind one with the same number of round communication, while preserving completeness and soundness properties}.
%As a result, one can \emph{upgrade} every verifiable CVQC protocol with blindness almost for free with the help of QFHE.
%
%Conceptually, we take a very different approach from previous ones  (e.g.,~\cite{FK17} as well as failed attempts in~\cite{mahadev_2018}) which use the blindness as the start point and then work to extend it with verifiability.
%
%Instead, our strategy is to simulate a (verifiable) CVQC protocol under QFHE per each message.
%To that end, we do require a special property of QFHE that the classical part of the ciphertext can be operated on separately from the quantum part, which is satisfied by the construction from~\cite{mahadev_qfhe}.
%Our construction makes a modular use of QFHE and only requires a minor technicality in the analysis, which will be explained below. 
%As a result, we obtain
\begin{theorem}[informal]
Assuming the QLWE assumption\footnote{By using Brakerski's QFHE, we only need to rely on the QLWE assumption with polynomial modulus in this theorem.}, there exists a protocol compiler that transforms any CVQC protocol $\Pi$ to a CVQC protocol $\Piblind$ that achieves blindness while preserves its round complexity, completeness, and soundness.
\end{theorem}

>>>>

Recently, the problem is resolved by a breakthrough result of Mahadev~\cite{FOCS:Mahadev18a}, who constructed the first Classical Verification of Quantum Computation (CVQC) protocol for $\BQP$, where an efficient classical ($\BPP$) verifier can interact with an efficient quantum ($\BQP$) prover to verify any $\BQP$ language. Soundness of Mahadev's protocol  is based on a widely recognized computational assumption that the learning with errors (LWE) problem~\cite{JACM:Regev09} is hard for $\BQP$ machines.
The technique invented therein has inspired many subsequent developments of CVQC protocols with improved parameters and functionality. For example, Mahadev's protocol has a large constant soundness error. The works of ~\cite{arXiv:AlaChiHun19,arXiv:ChiaChungYam19} use parallel repetition to achieve a negligible soundness error. As another example, the work of ~\cite{FOCS:GheVid19} extends Mahadev's techniques in an involved way to obtain a CVQC protocol with an additional blindness property.

@@

@@

We initiate the study of CVQC for quantum sampling problem, which we believe is highly desirable and natural for delegation of quantum computation. Due to the intrinsic randomness of quantum mechanics, the output from a quantum computation is randomized and described by a distribution. 
Thus, if a classical verifier want to utilize the full power of a quantum machine, the ability to get a verifiable sample from the quantum circuit's output distribution is desirable.  On a more concrete level, quantum algorithms like Shor's algorithm~\cite{Shor} has a significant quantum sampling component, and the recent quantum supremacy proposals (e.g.,~\cite{Boson, IQP, nature-google}) are built around sampling tasks, suggesting the importance of sampling in quantum computation.

It is worth noting that the difficulty of extending the delegation of decision problem to the delegation of sampling problems is quantum-specific. This is because 
there is a simple reduction from the delegation of \emph{classical} sampling problems to decision ones:  the verifier can sample and fix the random seed of the computation, 
which makes the computation deterministic. Then, the verifier can delegate the output of the computation bit-by-bit as decision problems. However, this derandomization trick does not work in the quantum setting due to its intrinsic randomness.

@@

\paragraph{Our Contribution.}
As the first step to formalize CVQC for quantum sampling problems, we consider the complexity class $\SampBQP$ introduced by Aaronson~\cite{aaronson_2013} as a natural class to capture efficiently computable quantum sampling problems. 
$\SampBQP$ consists of sampling problems $(D_x)_{x\in\zo^*}$ that can be approximately sampled by a $\BQP$ machine with a desired inverse polynomial error (See Section~\ref{sec:samp_definition} for the formal definition). We consider CVQC for a $\SampBQP$ problem $(D_x)_{x\in\zo^*}$ where a classical $\BPP$ verifier delegates the computation of a sample $z\leftarrow D_x$ for some input $x$ to a quantum $\BQP$ prover. Completeness requires that when the prover is honest, the verifier should accept with high probability and learn a correct sample $z\leftarrow D_x$. For soundness, intuitively, the verifier should not accept and output a sample with incorrect distribution  when interacting with a malicious prover. We formalize the soundness by a strong \emph{simulation-based} definition, (\Cref{dfn:stats-secure-proto-sampbqp}),
where we require that the joint distribution $(d,z)$ of the decision bit $d \in \set{\Acc, \Rej}$ and the output $z$ (which is $\bot$ when $d = \Rej$) is $\eps$-close (in  either statistical or computational sense) to an ``ideal distribution'' $(d,z_{ideal})$, where $z_{ideal}$ is sampled from the desired distribution $D_x$ when $d = \Acc$ and set to $\bot$ when $d = \Rej$.\footnote{This simulation-based formulation is analogous to the standard composable security definition for QKD.}

@@

@@

\paragraph{Overview of Mahadev's Protocol.}
Following~\cite{FOCS:Mahadev18a}, we define $\QPIP_{\tau}$ as classes of interactive proof systems between an (almost) classical verifier and a quantum prover, where the classical verifier has limited quantum computational capability, formalized as possessing $\tau$-qubit quantum memory.
A formal definition is given in Appendix~\ref{sec:qpip_def}.

@@

\paragraph{Overview of Our Construction.}
Following the blueprint of Mahadev's construction, our construction proceeds in the following two steps: 1. construct a $\QPIP_1$ protocol for $\SampBQP$ with required special property, and 2. compile the $\QPIP_1$ protocol using $\PiMeasure$ to get the desired $\QPIP_0$ protocol. The first step can be done by combining existing techniques from different contexts, whereas the second step is the main technical challenge.
At a high-level, the reason is the above-mentioned issue that the verifier does not learn the measurement outcome in the testing round. While this is not a problem for decision problems,
for sampling problems, the verifier needs to produce an output sample when accepts, but there seems to be no way to produce the output for the verifier without learning the measurement outcome.   
We discuss both steps in turn as follows.

@@

Clearly from the construction, the verifier now can output a sample when it decides to accept, and output a correct sample when interacting with an honest prover (completeness). The challenge is to show that $\PiSampZ$ is computationally sound. Since we are now in the computational setting, we cannot use the quantum de Finetti theorem as above which only holds in the information-theoretical setting. Furthermore, parallel repetition for computationally sound protocols are typically difficult to analyze, and known to not always work for protocols with four or more messages even in the classical setting~\cite{BIN97,PW12}.

 Parallel repetition of Mahadev's protocol for $\BQP$ has been analyzed before in ~\cite{arXiv:ChiaChungYam19, arXiv:AlaChiHun19}. However, the situation here is different. 
 For $\BQP$, the verifier simply chooses to run the Hadamard and testing rounds independently for each repetition.
 In contrast, our $\PiSampZ$ runs the Hadamard round in one repetition and runs the testing rounds in the rest. The reason is that in $\SampBQP$, as well as generically in sampling problems, there is no known approach to combine multiple samples to generate one sample with reduced error, i.e., there is no generic error reduction method for the sampling problem. 
In contrast, the error reduction for decision problems can be done with the majority vote. 
As a result, while the soundness error decreases exponentially for $\BQP$, as we see below (and also in the above $\QPIP_1$ protocols), for $\SampBQP$, $m$-fold repetition only decreases the error to $\poly(1/m)$.

To analyze the soundness of $\PiSampZ$, we use the \emph{partition lemma} developed in~\cite{arXiv:ChiaChungYam19} to analyze the prover's behavior while executing copies of $\PiMeasure$.\footnote{The analysis of \cite{arXiv:AlaChiHun19} is more tailored to the decision problems setting, and it is unclear how to extend it to sampling problems where there are multiple bits of output.} Intuitively, the partition lemma says that for any cheating prover and for each copy $i\in[m]$, there exist two efficient ``projectors" \footnote{Actually they are not projectors, but for the simplicity of this discussion let's assume they are.} $G_{0,i}$ and $G_{1,i}$ in the prover's internal space with $G_{0,i}+G_{1,i} \approx Id$. $G_{0,i}$ and $G_{1,i}$ splits up the prover's residual internal state after sending back his first message.
$G_{0,i}$ intuitively represents the subspace where the prover does not knows the answer to the testing round on the $i$-th copy, while $G_{1,i}$ represents the subspace where the prover does. Note that the prover is using a single internal space for all copies, and every $G_{0,i}$ and every $G_{1,i}$ is acting on this single internal space. 
By using this partition lemma iteratively, we can decompose the prover's internal state $\ket{\psi}$ into sum of subnormalized states.
First we apply it to the first copy, writing $\ket{\psi}=G_{0,1}\ket{\psi}+G_{1,1}\ket{\psi} \equiv \ket{\psi_0}+\ket{\psi_1}$.
The component $\ket{\psi_0}$ would then get rejected as long as the first copy is chosen as a testing round,
which occurs with pretty high probability.
More precisely, the output corresponding to $\ket{\psi_0}$ is $1/m$-close to the ideal distribution that just rejects all the time.
On the other hand, $\ket{\psi_1}$ is now binding on the first copy;
we now similarly apply the partition lemma of the second copy to $\ket{\psi_1}$.
We write $\ket{\psi_1}=G_{0,2}\ket{\psi_1}+G_{1,2}\ket{\psi_1}\equiv \ket{\psi_{10}}+\ket{\psi_{11}}$, and apply the same argument about $\ket{\psi_{10}}$ and $\ket{\psi_{11}}$.
We then continue to decompose $\ket{\psi_{11}}=\ket{\psi_{110}}+\ket{\psi_{111}}$ and so on, until we reach the last copy and obtain $\ket{\psi_{1^m}}$.
Intuitively, all the $\ket{\psi_{1\dots10}}$ terms will be rejected with high probability, while the $\ket{\psi_{1^m}}$ term represents the ``good" component where the prover knows the answer to every testing round and therefore has high accept probability. Therefore, $\ket{\psi_{1^m}}$ also satisfies some binding property,
so the verifier should obtain a measurement result of some state on the Hadamard round copy,
and the soundness of the $\QPIP_1$ protocol $\PiSamp$ follows.

@@

The intuitive analysis outlined above glosses over many technical details, and we substantiate this outline with full details in 
Section~\ref{sec:qpip0_all}.

@@

\begin{theorem}[informal]
Assuming the QLWE assumption\footnote{By using Brakerski's QFHE, we only need to rely on the QLWE assumption with polynomial modulus in this theorem.}, there exists a protocol compiler that transforms any CVQC protocol $\Pi$ to a CVQC protocol $\Piblind$ that achieves blindness while preserves its round complexity, completeness, and soundness.
\end{theorem}
